---
title: "SQL Connection"
format: html
---

```{r, message=FALSE}
library(DBI)
library(dplyr)
library(tidyverse)
```

For this project, we were given access to a database containing traffic stop data from across the United States. As someone who grew up in Oregon and now attends college in California, I was particularly interested in the three West Coast states—Oregon, California, and Washington. I wanted to examine the racial breakdown of traffic stops in each state and compare those figures to the actual population demographics. This comparison could reveal potential disparities in how different racial groups are stopped by law enforcement.

To begin the analysis, we connected to the database using secure credentials (not shown here to protect access).
```{r, echo=FALSE}
con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

Next, we analyzed the racial breakdown of traffic stops in each state. To do this, we examined the traffic stop data for Oregon, California, and Washington individually, then calculated the percentage of total stops attributed to each racial group within each state. This approach allowed us to standardize the data and make meaningful comparisons across states. The resulting table displays the racial composition of traffic stops side by side for all three states.

```{sql connection=con_traffic}
#| output.var: "west_coast_sql"
-- Oregon arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM or_statewide_2020_04_01) AS percentage,
  'Oregon' AS state
FROM or_statewide_2020_04_01
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

UNION ALL

-- Washington arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM wa_statewide_2020_04_01) AS percentage,
  'Washington' AS state
FROM wa_statewide_2020_04_01
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

UNION ALL

-- California arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM ca_statewide_2023_01_26) AS percentage,
  'California' AS state
FROM ca_statewide_2023_01_26
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

ORDER BY
  state,
  percentage DESC;
```

```{r}
west_coast_sql
```

I then created a side-by-side bar plot to visually compare the racial breakdown of traffic stops across the three states. The results showed that Oregon had the highest proportion of individuals identified as white in its traffic stop data. In contrast, California had the lowest proportion of white individuals and the highest proportion of Hispanic individuals—an expected trend given its large Hispanic population and proximity to the Mexican border. One particularly notable finding was the high percentage of traffic stops labeled as "unknown" in Washington. This suggests that either law enforcement officers in Washington are not required to record race/ethnicity during stops, or that there may be inconsistencies in how that information is collected.

```{r}
ggplot(west_coast_sql, aes(x = subject_race, y = percentage, fill = state)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Percentage of Each Race by State",
    x = "Race",
    y = "Percentage",
    fill = "State"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Next, I shifted focus from comparing the three states to examining the gap between the racial breakdown of traffic stops and the actual population demographics within each state. To do this, I incorporated U.S. Census data into the original table, aligning each racial group's share of traffic stops with their corresponding share of the state’s population. This allowed me to create a comparative chart that highlights potential disparities between who is being stopped and the overall demographic makeup of each state.

```{sql connection=con_traffic}
#| output.var: "combined_data"
-- Static population data
SELECT 
  subject_race, 
  NULL AS count_race, 
  percentage, 
  state
FROM (
  SELECT 'white' AS subject_race, 71.0 AS percentage, 'Oregon Population' AS state
  UNION ALL SELECT 'hispanic', 15.16, 'Oregon Population'
  UNION ALL SELECT 'asian', 4.4, 'Oregon Population'
  UNION ALL SELECT 'black', 1.8, 'Oregon Population'
  UNION ALL SELECT 'native american', 1.0, 'Oregon Population'
  UNION ALL SELECT 'pacific islander', 0.4, 'Oregon Population'
  UNION ALL SELECT 'two or more races', 6.2, 'Oregon Population'

  UNION ALL SELECT 'white', 63.0, 'Washington Population'
  UNION ALL SELECT 'hispanic', 14.6, 'Washington Population'
  UNION ALL SELECT 'asian', 9.3, 'Washington Population'
  UNION ALL SELECT 'black', 3.8, 'Washington Population'
  UNION ALL SELECT 'native american', 1.1, 'Washington Population'
  UNION ALL SELECT 'pacific islander', 0.7, 'Washington Population'
  UNION ALL SELECT 'two or more races', 7.5, 'Washington Population'

  UNION ALL SELECT 'white', 35.0, 'California Population'
  UNION ALL SELECT 'hispanic', 39.4, 'California Population'
  UNION ALL SELECT 'asian', 15.0, 'California Population'
  UNION ALL SELECT 'black', 5.0, 'California Population'
  UNION ALL SELECT 'native american', 0.8, 'California Population'
  UNION ALL SELECT 'pacific islander', 0.3, 'California Population'
  UNION ALL SELECT 'two or more races', 4.5, 'California Population'
) AS static_data

UNION ALL

-- Oregon arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM or_statewide_2020_04_01) AS percentage,
  'Oregon' AS state
FROM or_statewide_2020_04_01
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

UNION ALL

-- Washington arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM wa_statewide_2020_04_01) AS percentage,
  'Washington' AS state
FROM wa_statewide_2020_04_01
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

UNION ALL

-- California arrest data
SELECT
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END AS subject_race,
  COUNT(*) AS count_race,
  100.0 * COUNT(*) / (SELECT COUNT(*) FROM ca_statewide_2023_01_26) AS percentage,
  'California' AS state
FROM ca_statewide_2023_01_26
GROUP BY
  CASE WHEN subject_race IS NULL THEN 'unknown' ELSE subject_race END

ORDER BY
  state,
  percentage DESC;
```

Once the table was complete, we visualized the data using a faceted chart that placed traffic stop data on the left and population demographics on the right for each state. This layout allowed for easy side-by-side comparison. In California, we observed that white individuals made up a higher proportion of traffic stops than their share of the general population. The difference was most noticeable in the Hispanic category, where fewer Hispanic individuals were stopped relative to their proportion of the population. One possible explanation is that census data is self-reported, while officers may rely on visual assessments during traffic stops, potentially leading to underreporting or misclassification of ethnicity. A similar pattern appeared in Oregon, where discrepancies may also stem from differences in reporting methods. Interestingly, Washington was the only state where the percentage of white individuals was higher in the general population than in traffic stops. This reversal raises questions—either the large number of “unknown” entries in Washington's stop data is skewing the results, or there may be a difference in how traffic stops are being conducted or recorded in that state.

```{r}
ggplot(combined_data, aes(x = subject_race, y = percentage, fill = subject_race)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ state, nrow = 3, ncol = 2) +
  labs(
    title = "Original Racial Composition by State (from SQL Data)",
    x = "Race",
    y = "Percentage",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 12), 
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  ) +
  guides(fill = "none")
```

Overall, my goal was to examine how race is represented within traffic stop data both within and across the three West Coast states. The analysis revealed inconsistencies and limitations in the way racial data is collected during traffic stops, suggesting that the dataset may not be fully reliable. While some discrepancies between traffic stop data and census demographics could indicate potential bias, it is also possible that these gaps are due to poor or inconsistent data collection practices rather than intentional discrimination. This highlights the need for improved and standardized data collection procedures, so that future analyses can more accurately reflect demographic realities and support informed policy decisions.

```{r}
DBI::dbDisconnect(con_traffic)
```

**Refrences**

Pierson, Emma, Camelia Simoiu, Jan Overgoor, Sam Corbett-Davies, Daniel Jenson, Amy Shoemaker, Vignesh Ramachandran, et al. 2020. “A Large-Scale Analysis of Racial Disparities in Police Stops Across the United States.” Nature Human Behaviour, 1–10.

U.S. Census Bureau. 2023. QuickFacts: California; Oregon; Washington. U.S. Department of Commerce. Retrieved from https://www.census.gov/quickfacts/
