---
title: "The Office Line-Analysis"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
knitr:
  opts_chunk:
    warning: false
---

For this project, I chose to analyze a data set containing every line from the she "The Office". The data set was found at https://www.kaggle.com/datasets/fabriziocominetti/the-office-lines?resource=download. In this data set I wanted to look at how  the characters from my favorite show appeared through the show and how an iconic joke appeared as well.

```{r, message=FALSE}
#| echo: false
library(readr)
the_office_lines <- read_csv("the-office_lines.csv")
```

```{r, message=FALSE}
#| echo: false
library(tidyverse)
```

```{r, message=FALSE}
top_characters <- the_office_lines |>
  group_by(Character) |>
  summarize(total_count = n(), .groups = "drop") |>
  slice_max(total_count, n = 10)

lines_per_season <- the_office_lines |>
  filter(Character %in% top_characters$Character) |>
  group_by(Character, Season, add = TRUE) |>
  summarize(appearences = n(), .groups = "drop") |>
  group_by(Season) |>
  mutate(total_season = sum(appearences)) |>
  ungroup() |>
  mutate(percent = (appearences/total_season)*100)

ggplot(lines_per_season, aes(x = Season, y = percent, color = Character)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Season",
    y = "Percentage",
    title = "Line Breakdown Per Season"
  ) +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    plot.title = element_text(size = 18, face = "bold")
  )
```

The first thing that I analyzed is in the graph above, showing the distribution of lines characters have each season of the show. The graph shows the 10 characters with the highest total lines in the show and their share of each season of the office between the 10. As you can see Micheal dominates the share of lines through his time with the office, only slightly declining throughout. As he leaves the show following season 8, you see a jump in other characters roles as they needed to fill the gap that Michael left, such as Andy and Dwight. This graph shows the importance that Michael had to the show and how much it changed in its last two seasons.


```{r}
thats_what_she_said <- the_office_lines |>
  mutate(lower_lines = str_to_lower(Line)) |>
  mutate(quote = str_extract(lower_lines, "that’s what she said")) |>
  filter(!is.na(quote)) |>
  mutate(season_text = str_c(Season)) |>
  group_by(season_text) |>
  summarize(count = n(), .groups = "drop") |>
  add_row(count = 0, season_text = "1") |>
  add_row(count = 0, season_text = "8")

ggplot(thats_what_she_said, aes(x = season_text, y = count)) +
  geom_col() +
  labs(
    x = "Season",
    y = "Quantity of Jokes",
    title = "That's What She Said Jokes a Season"
  ) +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    plot.title = element_text(size = 18, face = "bold")
  )
```

The next thing I looked at focused on Michaels famous joke, "that's what she said". The chart above breaks down the usage in the show throughout the nine seasons. As you can see in the chart, they created the joke starting in season 2 and slowly declined in use as the show went on before Michael's departure following season season 7. There is a return in season 9, likely due to Michael's return for the finale.


```{r}
words_preceding_she_said <- the_office_lines |>
  mutate(lower_case = str_to_lower(Line)) |>
  mutate(char_following = str_extract(lower_case, "(?<=that’s what she said).")) |>
  filter(!is.na(char_following)) |>
  group_by(char_following) |>
  summarize(count = n())
  
ggplot(words_preceding_she_said, aes(x = char_following, y = count, text = char_following)) +
  geom_col() +
  labs(
    x = "Character following",
    y = "Appearences",
    title = 'Character following "Thats what she said"'
  ) +
  theme(
    axis.text.x = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold")
  )
```

Finally I was curious how this joke was typically written into the show. The chart above does this by breaking down which punctuation was used following the joke. This is able to tell us how the joke was presented. For example, we are able to figure out that more often than not, the joke is not written as Michael screaming out "That's what she said!", rather he delivers it less excited. You are also able to see that it is written in other ways as well such as the ', which is a line where Jim quotes the joke inside of a question. What we see is they often didn't direct the joke to be delivered with a lot of excitement, rather they left it up to the actor who was making the joke to decide how to deliver it.





