---
title: "The Office Line-Analysis"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
knitr:
  opts_chunk:
    warning: false
---

For this project, I chose to analyze a data set containing every line from "The Office". The data set was found [here](https://www.kaggle.com/datasets/fabriziocominetti/the-office-lines). In this data set I wanted to look at how the characters from my favorite show appeared through the show and how an iconic joke appeared as well.

```{r, message=FALSE}
#| echo: false
library(readr)
library(stringr)
the_office_lines <- read_csv("the-office_lines.csv")
```

```{r, message=FALSE}
#| echo: false
library(tidyverse)
```

```{r, message=FALSE}
top_characters <- the_office_lines |>
  filter(str_detect(Line, "\\[.*\\]")) |>
  group_by(Character) |>
  summarize(total_count = n(), .groups = "drop") |>
  slice_max(total_count, n = 10)

lines_per_season <- the_office_lines |>
  semi_join(top_characters, by = "Character") |>
  group_by(Character, Season, add = TRUE) |>
  summarize(appearences = n(), .groups = "drop") |>
  group_by(Season) |>
  mutate(total_season = sum(appearences)) |>
  ungroup() |>
  mutate(percent = (appearences/total_season)*100)

ggplot(lines_per_season, aes(x = Season, y = percent, color = Character)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    x = "Season",
    y = "Percentage",
    title = "Lines With Direction Breakdown Per Season"
  ) +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    plot.title = element_text(size = 18, face = "bold")
  )
```

The first thing I analyzed is the graph above, which shows the distribution of lines each character has per season in the show where they were given stage directions, designated with [ and ] somewhere in the line. The graph displays the 10 characters with the highest total lines containing these directions in the show and their share of these lines per season. As you can see, Michael dominates the share of lines containing stage directions throughout his time on The Office, with only a slight decline until a more significant drop after season 7. Other characters saw a slight increase following this drop, but for the most part, there was an equal split among the remaining characters in terms of the percentage of lines containing stage directions.


```{r}
thats_what_she_said <- the_office_lines |>
  mutate(lower_lines = str_to_lower(Line)) |>
  mutate(quote = str_extract(lower_lines, "that\\’?s what she said")) |>
  filter(!is.na(quote)) |>
  mutate(season_text = str_c(Season)) |>
  group_by(season_text) |>
  summarize(count = n(), .groups = "drop") |>
  add_row(count = 0, season_text = "1") |>
  add_row(count = 0, season_text = "8")

ggplot(thats_what_she_said, aes(x = season_text, y = count)) +
  geom_col() +
  labs(
    x = "Season",
    y = "Quantity of Jokes",
    title = "That's What She Said Jokes a Season"
  ) +
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    plot.title = element_text(size = 18, face = "bold")
  )
```

The next thing I looked at focused on Michael's famous joke, "that's what she said". The chart above breaks down the usage in the show throughout the nine seasons. As you can see in the chart, they created the joke starting in season 2 and slowly declined in use as the show went on before Michael's departure following season season 7. There is a return in season 9, likely due to Michael's return for the finale.


```{r}
words_preceding_she_said <- the_office_lines |>
  mutate(lower_case = str_to_lower(Line)) |>
  mutate(char_following = str_extract(lower_case, "(?<=that’s what she said).")) |>
  filter(!is.na(char_following)) |>
  group_by(char_following) |>
  summarize(count = n())
  
ggplot(words_preceding_she_said, aes(x = char_following, y = count, text = char_following)) +
  geom_col() +
  labs(
    x = "Character following",
    y = "Appearences",
    title = 'Character following "Thats what she said"'
  ) +
  theme(
    axis.text.x = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold")
  )
```

Finally I was curious how this joke was typically written into the show. The chart above does this by breaking down which punctuation was used following the joke. This is able to tell us how the joke was presented. For example, we are able to figure out that more often than not, the joke is not written as Michael screaming out "That's what she said!", rather he delivers it less excited. You are also able to see that it is written in other ways as well such as the ', which is a line where Jim quotes the joke inside of a question. What we see is they often didn't direct the joke to be delivered with a lot of excitement, rather they left it up to the actor who was making the joke to decide how to deliver it.





